Bee visitation, pollination services, and plant yield in commodity and hybrid seed canola

### 2022-08-15 Manuscript/main.tex
Worked on Discussion

%REVIEW THIS
The limited pollen-transfering behaviour of HBs was also reflected in the pollen counts, as deposition strongly increased with LCB visitation, but not with HB visitation.
LCBs tend to switch between male and female flowers more frequently than HBs do \citep{waytesMsc}, transport more viable pollen between flowers \citep{parker2015}, and tend not to side-work canola flowers \citep{soroka2001}, which may explain their increased pollination efficacy.
Pollen deposition in seed fields also decreased with distance from the edge of the field, and was lower in the centre of the female bays, but this was not explained solely by LCB abundance (Figure \ref{fig:allSEM}).
During foraging bouts, LCBs may travel further between flowers than HBs do \citep{brunet2019}, but pollen carryover from the edge to the centre of the bay female is likely reduced, as LCBs likely visit the edge of the female bay before venturing into the centre \citep{thomson1986, pinnisch1990}.
Pollen counts decreased with distance into the field (independent of visitation), suggests that pollinators are less efficient at transfering pollen.
Foraging \textit{Bombus} take shorter trips between flowers under nectar-rich conditions (area-restricted search, \citealp{pyke1978b, heinrich1979}), suggesting that HBs and LCBs may operate similarly at the centre of seed fields, causing lower pollen deposition at the centre of the field where competition is lower.
% This may be due to lower levels of competition at the centre of the field, leading to a higher nectar (or pollen) standing crop. %LH: more likely that bees at centre of F bays carry less pollen
%LCBs are extremely fast fliers, and are difficult to follow after they leave an inflorescence, meaning that tracking their movements during an entire foraging bout would be almost impossible.
% Visits by HB foragers were almost twice as high in the centre of the female bay, but pollen deposition was about 40\% lower (  Figure \ref{fig:allPollen}), which correspondingly is 
% HBs in seed fields tended to visit more in the centre of the female bay, even close to the edge of the field, yet pollen deposition was still about 40\% lower (Figure \ref{fig:allPollen}).
% While we found that LCBs also tend to avoid the centre of the female bays, and decrease with distance into the field, this extra dependence could have been caused by the fact that the visitation data represents a short ``snapshot'' of the pollen deposition process.
% For example, canola flowers are receptive for 1-2 days before the petals dehisce, while the surveys only recorded 5 or 10 minutes of visitation, meaning we may have not captured the deposition process on the time scale that it occurs, and the extra variance is being accounted for by distance from field edge and the centre of the female bay.
% Leafcutter visitation can be extremely sporadic, so the distance and bay position measurements capture differences in the average visitation process better than 10 minute measurements.
% Pollen deposition also likely changes over the course of the season, as flower availability decreases.

@@

Plant size increased pod set in commodity canola, but both plant size and pollen increased pod set in seed canola, suggesting that both pollen deposition and plant resources constrain pod production.
This is similar to the findings of \citet{mesquida1981} and \citet{steffan2003}, who found that pod set in male-sterile plants responded positively to visitation, while male-fertile plants had no response (but see \citealp{adegas1992}).
Both pod set and seeds per pod can increase with extra pollination \citep{jauker2008,sabbahi2005,sabbahi2006,duran2010}, so the low influence of pollination in commodity fields may be due to a high overall level of pollen deposition (Figure \ref{fig:allPollen}).
There was no additional effect of HB visitation on pod set, meaning that that HB visitation did not cause more pollen deposition or improve its quality through increased outcrossing (as in \citealt{rosa2011}), but this may depend on the variety \citep{adamidis2019}.
Flowers per plant was reduced by pod set, implying that high pod set suppresses further flower production or increases competition among pods for plant resources.
The first scenario is more likely, as \citet{sabbahi2006} and \citet{mesquida1981} both found that canola plants compensated for experimental removal of flowers by increasing branch and flower production, until the plant reached about 170 pods.
There were also effects of distance on pod set independent of pollen: pod set decreased with distance from the edge of the field, distance to LCB shelters, and was lower at the centre of the female bays.
\citet{mesquida1978} also found that pod set in male-sterile canola declined quickly with distance from the male-fertile plants (due to wind pollination), but the effect we found is likely related to the earlier effect of lower pollen carryover at the centre of the female bays, and possibly the centre of the field.
Finally, the effect size of pollen deposition (Z=3.1) was much smaller than that of plant size (Z=19.9), indicating that the factors controlling plant size (e.g. fertilizer application, soil quality) likely constrain pod production more strongly than pollination alone \citep{marini2015}.

@@

Plant size increased seeds per pod in both crop types, but the number of seeds per pod in seed canola was also increased by pollen deposition and pod set. 
Pollen deposition had a smaller effect than plant size (Z=2.3 vs. 5.6), again suggesting that plant resources limit seed production more strongly than pollen deposition in seed canola \citep{marini2015}.
\citet{mesquida1982} also found that pod production, rather than seeds per pod, was the main variable controlled by pollen deposition in seed canola.
Position in the female bay influenced pollen deposition, both by wind and insects, indicating that the decrease in seed number at the centre of the bay is ultimately caused by reduced pollen deposition from lower visitation \citep{mesquida1981,mesquida1982}.

Seed size was influenced by plant size in seed canola, but not commodity canola, and was not influenced by pollen or visitation in either crop \citep{riffkin2012}.
Unexpectedly, plant density had a \emph{positive} effect on seed size in both crop types, which we consider could have been caused by two different processes:
plant density could be positively related to resource availability if plant survival is higher in plots with better growing conditions (i.e. Berkson's paradox \citep{snoep2014}) or plant density may improve the microclimate conditions of the canola stand, reducing heat or desiccation stress on individual plants.
We consider the first process more likely, as planting density under uniform conditions typically has little effect on seed size \citep{Angadi2003}.

>>>>

%REVIEW THIS
The limited pollen-transferring behaviour of HBs was also reflected in the pollen counts, as deposition strongly increased with LCB visitation, but not with HB visitation.
LCBs tend to switch between male and female flowers more frequently than HBs do \citep{waytesMsc}, transport more viable pollen between flowers \citep{parker2015}, and tend not to side-work canola flowers \citep{soroka2001}, which may explain their increased pollination efficacy.
Pollen deposition in seed fields also decreased with distance from the edge of the field, and was lower in the centre of the female bays, but this was not explained solely by LCB abundance (Figure \ref{fig:allSEM}).
During foraging bouts, LCBs may travel further between flowers than HBs do \citep{brunet2019}, but pollen carryover from the edge to the centre of the bay female is likely reduced, as LCBs likely visit the edge of the female bay before venturing into the centre \citep{thomson1986, pinnisch1990}.
Pollen counts decreased with distance into the field (independent of visitation), suggests that pollinators are less efficient at transfering pollen.
Foraging \textit{Bombus} take shorter trips between flowers under nectar-rich conditions (area-restricted search, \citealp{pyke1978b, heinrich1979}), suggesting that HBs and LCBs may operate similarly at the centre of seed fields, causing lower pollen deposition at the centre of the field where competition is lower.
% This may be due to lower levels of competition at the centre of the field, leading to a higher nectar (or pollen) standing crop. %LH: more likely that bees at centre of F bays carry less pollen
%LCBs are extremely fast fliers, and are difficult to follow after they leave an inflorescence, meaning that tracking their movements during an entire foraging bout would be almost impossible.
% Visits by HB foragers were almost twice as high in the centre of the female bay, but pollen deposition was about 40\% lower (  Figure \ref{fig:allPollen}), which correspondingly is 
% HBs in seed fields tended to visit more in the centre of the female bay, even close to the edge of the field, yet pollen deposition was still about 40\% lower (Figure \ref{fig:allPollen}).
% While we found that LCBs also tend to avoid the centre of the female bays, and decrease with distance into the field, this extra dependence could have been caused by the fact that the visitation data represents a short ``snapshot'' of the pollen deposition process.
% For example, canola flowers are receptive for 1-2 days before the petals dehisce, while the surveys only recorded 5 or 10 minutes of visitation, meaning we may have not captured the deposition process on the time scale that it occurs, and the extra variance is being accounted for by distance from field edge and the centre of the female bay.
% Leafcutter visitation can be extremely sporadic, so the distance and bay position measurements capture differences in the average visitation process better than 10 minute measurements.
% Pollen deposition also likely changes over the course of the season, as flower availability decreases.

@@

Plant size increased the rate of pod set in commodity canola, but both plant size and pollen increased pod set in seed canola, suggesting that both pollen deposition and plant resources constrain pod production.
This is similar to the findings of \citet{mesquida1981} and \citet{steffan2003}, who found that pod set in male-sterile plants responded positively to visitation, while male-fertile plants had no response (but see \citealp{adegas1992}).
Both pod set and seeds per pod can increase with extra pollination \citep{jauker2008,sabbahi2005,sabbahi2006,duran2010}, so the low influence of pollination in commodity fields may be due to a high overall level of pollen deposition (Figure \ref{fig:allPollen}).
There was no additional effect of HB visitation on pod set, meaning that that HB visitation did not cause more pollen deposition or improve its quality through increased outcrossing (as in \citealt{rosa2011}), but this may depend on the variety \citep{adamidis2019}.
Flowers per plant was reduced by pod set, implying that high pod set suppresses further flower production or increases competition among pods for plant resources.
The first scenario is more likely, as \citet{sabbahi2006} and \citet{mesquida1981} both found that canola plants compensated for experimental removal of flowers by increasing branch and flower production, until the plant reached about 170 pods.
There were also effects of distance on pod set independent of pollen: pod set decreased with distance from the edge of the field, distance to LCB shelters, and was lower at the centre of the female bays.
\citet{mesquida1978} also found that pod set in male-sterile canola declined quickly with distance from the male-fertile plants (due to wind pollination), but the effect we found is likely related to the earlier effect of lower pollen carryover at the centre of the female bays, and possibly the centre of the field.
Finally, the effect size of pollen deposition (Z=3.1) was much smaller than that of plant size (Z=19.9), indicating that the factors controlling plant size (e.g. fertilizer application, soil quality) likely constrain pod production more strongly than pollination alone \citep{marini2015}.

@@

Extra pollination generally results in canola plants producing more, smaller-sized seeds (see review in \citealt{ouvrard2019}, but see also \citealtt{koltowski2005}), whereby the plant allocates the same number of resources to a greater number of fertilized ovules.
We found mixed support across both crop types: seed size and number in commodity canola was mainly influenced by plant size, while seed canola was influenced by both plant size and pollen deposition.   
Seeds per pod was increased by plant size in both crop types, suggesting greater rates of ovule maturation in the presence of greater plant resources \citep{bouttier1992}.

Pollen deposition had a smaller effect than plant size (Z = 2.3 vs. 5.6), again suggesting that plant resources limit seed production more strongly than pollen deposition in seed canola \citep{marini2015}.

In contrast, we found that seed size was influenced by plant size in seed canola alone, and was not influenced by pollen or visitation in either crop.

While we d

Unexpectedly, plant density had a \emph{positive} effect on seed size in both crop types, which we consider could have been caused by two different processes:
plant density could be positively related to resource availability if plant survival is higher in plots with better growing conditions (i.e. Berkson's paradox, \citep{snoep2014}) or plant density may improve the microclimate conditions of the canola stand, reducing heat or desiccation stress on individual plants.
We consider the first process more likely, as planting density under uniform conditions typically has little effect on seed size \citep{Angadi2003}.

% The decrease in seed number at the centre of the bay may ultimately be caused by reduced pollen deposition from lower visitation \citep{mesquida1981,mesquida1982}.
